Long-form recording of infant body position in the home using wearable inertial sensors

### 2023-03-15 manuscript.Rmd
more discussion

How did classification in the current investigation compare to prior work [@FranchakScott2021; @AiraksinenRasanen2020; @AiraksinenGallen2022]? Across body positions, median accuracy was `r metrics_lookup["Overall Accuracy", "median_Group"]*100`% for group models (Kappa = `r metrics_lookup["Kappa", "median_Group"]*100`) and `r metrics_lookup["Overall Accuracy", "median_Individual"]`% for individual models (Kappa = `r metrics_lookup["Kappa", "median_Individual"]`) during the proximal comparison. Although these accuracy and kappa values are slightly lower than in past work (accuracy 95-98%, kappas .93-.95), we note that those past values were obtained under ideal circumstances. A researcher applied the sensors in the lab [@FranchakScott2021; @AiraksinenRasanen2020] or home [@AiraksinenGallen2022] and set the stage for how infants and caregivers should interact. For example, both @FranchakScott2021 and @AiraksinenGallen2022 instructed the caregivers that the recorded sessions should mimic a "playtime", which would encourage more common activities that would be easier to classify (e.g., crawling and sitting on the floor) and discourage idiosyncratic and potentially more challenging activities (e.g., positioning in infant-specific furniture like exersaucers and walkers, sitting or lying for long periods while eating or nursing). Examining the timelines in Figure \@ref(fig:timelines) shows how much variability there is between and within sessions in the earlier parts of the recording that were used for training and testing. With the exception of the brief guided session at the start of the recording (which ensured that every infant spent at least a short amount of time in each body position), the majority of the recorded behavior was truly natural and "in the wild".

@@

To put this in context, the reliability of human coders on the body position code was ~90%, putting model-predicted accuracy on a similar level to human coders. Acceptable inter-rater reliability in developmental psychology studies of motor behavior is greater than 80%; both group and individual model results exceed that standard for most recorded sessions. Another comparison we can make is to the accuracy of the LENA&reg;'s automated metrics. Correlations between human and model frequencies were *r* = .698 for adult word count and *r* = .649 for child word count [@CristiaLavechin2020]. Even including outliers, correlations between human and model-predicted body position were *r* = `r group_agree_cor` (group) and *r* = `r split_agree_cor` (individual), easily surpassing the LENA&reg; device's most reliable metrics. 
<!-- Reliability standard? -->

Outlier participants give us more confidence that it actually works as intended. Disagreements limited to some edge cases, not all events/classes. Even w/ outliers (which are they on timelines), most of the rest of the data seems typical for the age group. Some outlier-ish data in age comparisons, but nothing out of the ordinary (hard to know if it's the model or the behavior)

Nice to have video to confirm. Best practice in LENA is to still do some transcription yourself to see what cases it works. Still helpful to have other sources to triangulate against. Maybe don't use it for a holding detector.

@@

See how other aspects of behavior align w/ movement in real time. See how new motor skills relate to chg in daily behavior.

>>>>

How did classification in the current investigation compare to prior work [@FranchakScott2021; @AiraksinenRasanen2020; @AiraksinenGallen2022]? Across body positions, median accuracy was `r metrics_lookup["Overall Accuracy", "median_Group"]*100`% for group models (Kappa = `r metrics_lookup["Kappa", "median_Group"]`) and `r metrics_lookup["Overall Accuracy", "median_Individual"]*100`% for individual models (Kappa = `r metrics_lookup["Kappa", "median_Individual"]`) during the proximal comparison. Although these accuracy and kappa values are slightly lower than in past work (accuracy 95-98%, kappas .93-.95), we note that those past values were obtained under ideal circumstances. A researcher applied the sensors in the lab [@FranchakScott2021; @AiraksinenRasanen2020] or home [@AiraksinenGallen2022] and set the stage for how infants and caregivers should interact. For example, both @FranchakScott2021 and @AiraksinenGallen2022 instructed the caregivers that the recorded sessions should mimic a "playtime", which would encourage more common activities that would be easier to classify (e.g., crawling and sitting on the floor) and discourage idiosyncratic and potentially more challenging activities (e.g., positioning in infant-specific furniture like exersaucers and walkers, sitting or lying for long periods while eating or nursing). Examining the timelines in Figure \@ref(fig:timelines) shows how much variability there is between and within sessions in the earlier parts of the recording that were used for training and testing. With the exception of the brief guided session at the start of the recording (which ensured that every infant spent at least a short amount of time in each body position), the majority of the recorded behavior was truly natural and "in the wild".

@@

To put this in context, the reliability of human coders on the body position code was ~90%, putting model-predicted accuracy on a similar level to human coders. Acceptable inter-rater reliability in developmental psychology studies of motor behavior is greater than 80%; both group and individual model results exceed that standard for most recorded sessions. Another comparison we can make is to the accuracy of the LENA&reg;'s automated metrics. Correlations between human and model frequencies were *r* = .698 for adult word count and *r* = .649 for child word count [@CristiaLavechin2020]. Even including outliers, correlations between human and model-predicted body position were *r* = `r group_agree_cor` (group) and *r* = `r split_agree_cor` (individual), stronger than the LENA&reg; automated counts. Note, however, that the LENA&reg; validation study used independently-collected data, whereas we are measuring performance for infants whose earlier data comprised the training set. More work is needed to know how well group models would predict a different set of infants' behavior in a different environmental and/or cultural context. 
<!-- Reliability standard? -->

In their validation study, @CristiaLavechin2020 make the recommendation that researchers using LENA&reg; manually annotate a portion of data collected for their particular use case to ensure that automated LENA&reg; metrics are reliable for the participants and environment pertinent to the research question. Similarly, collecting some video data for each participant in future long-form motor behavior recordings provides a way to check reliability and identify errant predictions. Although predictions were accurate for most sessions, we noted two outliers for whom model predictions failed to match human annotations. Fortunately, we could reference video recordings to understand why the errors occurred. Idiosyncrasies in activities and device use make it challenging to decide *a priori* whether odd looking classifications are possible. For example, younger infant \#8 on Figure \@ref(fig:timelines) spent an unusual amount of time upright at an age where infants cannot yet stand. Inspecting the video revealed that the infant spent long periods of time suspended in a jumper that supported their body in an upright position. Although it is not feasible to collect large amounts of video data to check model predictions, interviewing caregivers about common activities and devices may provide a way to understand unexpected predictions.

Counterintuitively, the two outlier participants increased our confidence that the body position classification method works as intended. At first glance, seeing low agreement rates for two sessions would suggest that the models perform poorly at predicting some *infants'* behavior. Instead, we found that for those infants, confusion was restricted to two positions during a single, long-lasting event and the other three body positions continued to be classified correctly. In other words, the models failed to predict a particular *event* for those two infants. Unfortunately, those events lasted a long time. In the case of confusion between supine and sitting for the infant who was reclined, the confusion likely stems from the difficult of categorizing behaviors that do not always fit neatly into discrete classes. Even to a human observer, there are often "edge" cases when scoring infant behavior, which is why inter-rater reliability is never perfect.

@@

See how other aspects of behavior align w/ movement in real time. See how new motor skills relate to chg in daily behavior. 
LINK TO AGE DIFFS. Some outlier-ish data in age comparisons, but nothing out of the ordinary (hard to know if it's the model or the behavior)

### 2023-03-15 manuscript.tex
more discussion

How did classification in the current investigation compare to prior work (Airaksinen et al., 2022, 2020; Franchak et al., 2021)? Across body positions, median accuracy was 89.40\% for group models (Kappa = 76.80) and 0.93\% for individual models (Kappa = 0.85) during the proximal comparison. Although these accuracy and kappa values are slightly lower than in past work (accuracy 95-98\%, kappas .93-.95), we note that those past values were obtained under ideal circumstances. A researcher applied the sensors in the lab (Airaksinen et al., 2020; Franchak et al., 2021) or home (Airaksinen et al., 2022) and set the stage for how infants and caregivers should interact. For example, both Franchak et al. (2021) and Airaksinen et al. (2022) instructed the caregivers that the recorded sessions should mimic a ``playtime'', which would encourage more common activities that would be easier to classify (e.g., crawling and sitting on the floor) and discourage idiosyncratic and potentially more challenging activities (e.g., positioning in infant-specific furniture like exersaucers and walkers, sitting or lying for long periods while eating or nursing). Examining the timelines in Figure \ref{fig:timelines} shows how much variability there is between and within sessions in the earlier parts of the recording that were used for training and testing. With the exception of the brief guided session at the start of the recording (which ensured that every infant spent at least a short amount of time in each body position), the majority of the recorded behavior was truly natural and ``in the wild''.

@@

To put this in context, the reliability of human coders on the body position code was \textasciitilde90\%, putting model-predicted accuracy on a similar level to human coders. Acceptable inter-rater reliability in developmental psychology studies of motor behavior is greater than 80\%; both group and individual model results exceed that standard for most recorded sessions. Another comparison we can make is to the accuracy of the LENA®'s automated metrics. Correlations between human and model frequencies were \emph{r} = .698 for adult word count and \emph{r} = .649 for child word count (Cristia et al., 2020). Even including outliers, correlations between human and model-predicted body position were \emph{r} = 0.80 (group) and \emph{r} = 0.94 (individual), easily surpassing the LENA® device's most reliable metrics.

Outlier participants give us more confidence that it actually works as intended. Disagreements limited to some edge cases, not all events/classes. Even w/ outliers (which are they on timelines), most of the rest of the data seems typical for the age group. Some outlier-ish data in age comparisons, but nothing out of the ordinary (hard to know if it's the model or the behavior)

Nice to have video to confirm. Best practice in LENA is to still do some transcription yourself to see what cases it works. Still helpful to have other sources to triangulate against. Maybe don't use it for a holding detector.

>>>>

How did classification in the current investigation compare to prior work (Airaksinen et al., 2022, 2020; Franchak et al., 2021)? Across body positions, median accuracy was 89.40\% for group models (Kappa = 0.77) and 93.30\% for individual models (Kappa = 0.85) during the proximal comparison. Although these accuracy and kappa values are slightly lower than in past work (accuracy 95-98\%, kappas .93-.95), we note that those past values were obtained under ideal circumstances. A researcher applied the sensors in the lab (Airaksinen et al., 2020; Franchak et al., 2021) or home (Airaksinen et al., 2022) and set the stage for how infants and caregivers should interact. For example, both Franchak et al. (2021) and Airaksinen et al. (2022) instructed the caregivers that the recorded sessions should mimic a ``playtime'', which would encourage more common activities that would be easier to classify (e.g., crawling and sitting on the floor) and discourage idiosyncratic and potentially more challenging activities (e.g., positioning in infant-specific furniture like exersaucers and walkers, sitting or lying for long periods while eating or nursing). Examining the timelines in Figure \ref{fig:timelines} shows how much variability there is between and within sessions in the earlier parts of the recording that were used for training and testing. With the exception of the brief guided session at the start of the recording (which ensured that every infant spent at least a short amount of time in each body position), the majority of the recorded behavior was truly natural and ``in the wild''.

@@

To put this in context, the reliability of human coders on the body position code was \textasciitilde90\%, putting model-predicted accuracy on a similar level to human coders. Acceptable inter-rater reliability in developmental psychology studies of motor behavior is greater than 80\%; both group and individual model results exceed that standard for most recorded sessions. Another comparison we can make is to the accuracy of the LENA®`s automated metrics. Correlations between human and model frequencies were \emph{r} = .698 for adult word count and \emph{r} = .649 for child word count (Cristia et al., 2020). Even including outliers, correlations between human and model-predicted body position were \emph{r} = 0.80 (group) and \emph{r} = 0.94 (individual), stronger than the LENA® automated counts. Note, however, that the LENA® validation study used independently-collected data, whereas we are measuring performance for infants whose earlier data comprised the training set. More work is needed to know how well group models would predict a different set of infants' behavior in a different environmental and/or cultural context.

In their validation study, Cristia et al. (2020) make the recommendation that researchers using LENA® manually annotate a portion of data collected for their particular use case to ensure that automated LENA® metrics are reliable for the participants and environment pertinent to the research question. Similarly, collecting some video data for each participant in future long-form motor behavior recordings provides a way to check reliability and identify errant predictions. Although predictions were accurate for most sessions, we noted two outliers for whom model predictions failed to match human annotations. Fortunately, we could reference video recordings to understand why the errors occurred. Idiosyncrasies in activities and device use make it challenging to decide \emph{a priori} whether odd looking classifications are possible. For example, younger infant \#8 on Figure \ref{fig:timelines} spent an unusual amount of time upright at an age where infants cannot yet stand. Inspecting the video revealed that the infant spent long periods of time suspended in a jumper that supported their body in an upright position. Although it is not feasible to collect large amounts of video data to check model predictions, interviewing caregivers about common activities and devices may provide a way to understand unexpected predictions.

Counterintuitively, the two outlier participants increased our confidence that the body position classification method works as intended. At first glance, seeing low agreement rates for two sessions would suggest that the models perform poorly at predicting some \emph{infants'} behavior. Instead, we found that for those infants, confusion was restricted to two positions during a single, long-lasting event and the other three body positions continued to be classified correctly. In other words, the models failed to predict a particular \emph{event} for those two infants. Unfortunately, those events lasted a long time. In the case of confusion between supine and sitting for the infant who was reclined, the confusion likely stems from the difficult of categorizing behaviors that do not always fit neatly into discrete classes. Even to a human observer, there are often ``edge'' cases when scoring infant behavior, which is why inter-rater reliability is never perfect.
